Supporting Information for ``The dynamics of starvation and recovery''

### 2017-10-18 NatureComm_REVISION/YKR_supp-Revision-1.tex
new revision

%\usepackage{scicite}
%\usepackage{times}
%\usepackage{graphicx}
%\usepackage{verbatim}
%\usepackage{subfigure}
%\usepackage{tikz}
%\usepackage{array}
%\usepackage{tabularx}
%\usepackage{multirow}
%\usepackage{multicol}
%\usepackage{multibox}
%\usepackage{rotating}
%\usepackage{layout}
%\usepackage{epstopdf}
%\usepackage{afterpage}
%\usepackage[left]{lineno}

%% PNAStmpl.tex
%% Template file to use for PNAS articles prepared in LaTeX
%% Version: Apr 14, 2008

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% BASIC CLASS FILE

\documentclass{pnastwo}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% OPTIONAL GRAPHICS STYLE FILE

\usepackage[pdftex]{graphicx}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% OPTIONAL POSTSCRIPT FONT FILES

\usepackage{pnastwof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% ADDITIONAL OPTIONAL STYLE FILES

%\usepackage{amssymb,amsfonts,amsmath}
\usepackage{amsmath}
\usepackage{amssymb}

\usepackage[font=large]{caption}
\usepackage{subfigure}
\usepackage{epstopdf}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{array}
\usepackage{tabularx}
\usepackage{multirow}
\usepackage{color}
%\usepackage{multibox}
\usepackage{rotating}
\usepackage{color}
%\usepackage{lineno}
\usepackage[left]{lineno}
\usepackage[comma,sort&compress]{natbib}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% OPTIONAL MACRO FILES

\bibliographystyle{pnas}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\sid}[1]{\textcolor{red}{\bf [#1]}}

%% Don't type in anything in the following section:
%%%%%%%%%%%%
%% For PNAS Only:
\contributor{Submitted to Proceedings
of the National Academy of Sciences of the United States of America}
\url{www.pnas.org/cgi/doi/10.1073/pnas.0709640104}
\copyrightyear{2011}
\issuedate{Issue Date}
\volume{Volume}
\issuenumber{Issue Number}
%%%%%%%%%%%%

@@

\linenumbers
\setlength\linenumbersep{3pt}

%\title{Ecological and evolutionary implications of starvation and body size}

%\title{Survival of the Stable: Dynamic Consequences of Starvation and Reproduction}

%\title{Size Constraints on Starvation \& Reproduction Dynamics}

%\title{Eco-evolutionary Consequences of Starvation and Reproduction: Survival
%of the Stable}

\title{Supporting Information for ``The dynamics of starvation and recovery''}%: Eco-evolutionary feedbacks}

\author
{Justin D. Yeakel\affil{1}{School of Natural Science, University of California Merced, Merced, CA} \affil{2}{The Santa Fe Institute, Santa Fe, NM}\affil{4}{Contributed equally}\affil{5}{To whom correspondence should be addressed: jdyeakel@gmail.com}, Christopher P. Kempes \affil{2}{The Santa Fe Institute, Santa Fe, NM}\affil{4}{Contributed equally}, \and Sidney Redner \affil{2}{The Santa Fe Institute, Santa Fe, NM}\affil{3}{Department of Physics, Boston University, Boston MA}\affil{4}{Contributed equally}
}

\contributor{Submitted to Proceedings of the National Academy of Sciences
of the United States of America}

@@

\begin{article}

\section*{Mechanisms of Starvation and Recovery}
Our overall goal is to understand the dynamics of starvation, recovery, reproduction, and resource competition, where our framework partitions starvation and reproduction into two classes of the consumer: a full class that is able to reproduce and a hungry class that experiences mortality at a given rate and is unable to reproduce. For the dynamics of growth, reproduction, and resource consumption, past efforts have combined the overall metabolic rate as dictated by body size with a growth rate that is dependent on resource abundance and, in turn, dictates resource consumption (see Refs. \cite{Kempes:2012hy,kempes2014morphological} for a brief review of this perspective). This approach has been used to understand a range of phenomena including a derivation of ontogenetic growth curves from a partitioning of metabolism into maintenance and biosynthesis (e.g. \cite{West:2001bv,moses2008rmo,hou,Kempes:2012hy}) and predictions for the steady-state resource abundance in communities of cells \cite{kempes2014morphological}. Here we leverage these mechanisms, combined with several additional concepts, to define our Nutritional State Model (NSM).

We consider the following generalized set of explicit dynamics for starvation, recovery, reproduction, and resource growth and consumption
\begin{align}
\begin{split}
\dot{F_{d}} &= \lambda_{max} F_{d} + \rho_{max}R_{d}H_{d}/k - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho_{max}R_{d} H_{d}/k - \mu H_{d},  \\
\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) -\\
& \left[\left(\frac{\rho_{max}R_{d}}{Y_{H}k}+P_{H}\right)H_{d}+\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right)F_{d}\right].
\end{split}
\end{align}
where each term has a mechanistic meaning that we detail below (we will denote the dimensional equations with $_{d}$ before introducing the non-dimensional form that is presented in the main text). In the above equations $Y$ represents the yield coefficient (e.g., Refs. \citenum{pirt,Heijnen}) which is the quantity of resources required to build a unit of organism (gram of mammal produced per gram of resource consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism). If we pick $F_{d}$ and $H_{d}$ to have units of (g organisms $\cdot$ m$^{-2}$), then all of the terms of $\dot{R_{d}}$, such as $\frac{\rho\left(R_{d}\right)}{Y}H_{d}$, have units of (g resource $\cdot$ m$^{-2}$ $\cdot$ s$^{-1}$) which are the units of net primary productivity (NPP), a natural choice for $\dot{R_{d}}$. This choice also gives $R_{d}$ as (g $\cdot$ m$^{-2}$) which is also a natural unit and is simply the biomass density. In this system of units $\alpha$ (s$^{-1}$) is the specific growth rate of $R_{d}$, and $C$ is the carrying capacity, or maximum density, of $R_{d}$ in a particular environment, and $k$ is the half-saturation constant (half the density of resources that would lead to maximum growth).
%(note this is not fully explicit because I don't know how to deal with the response of $\sigma$ to resources, although I have an idea for a derivation which may be necessary given the following approximations)

@@

We can formally non-dimensionalize this system by choosing the general rescaling of $F=fF_{d}$, $H=fH_{d}$, $R=qR_{d}$, $t=st_{d}$, in which case our system of equations becomes
%(ignoring the $\sigma (1-R)F$ terms which I don't have a dimensional form for yet),:
\begin{align}
\begin{split}
\dot{F} &= \frac{1}{s}\left[\lambda_{max} F + \rho_{max}\frac{R}{qk}H - \sigma \left(1-\frac{R}{qC}\right)F\right],  \\
\dot{H} &= \frac{1}{s}\left[\sigma \left(1-\frac{R}{qC}\right)F - \rho_{max}\frac{R}{qk} H - \mu H\right],  \\
\dot{R} &= \frac{1}{s}\left[\alpha R\left(1-\frac{R}{qC}\right) -\frac{q}{f}\left[\left(\frac{\rho_{max}R}{Y_{H}kq}+P_{H}\right)H+\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right) F\right]\right].
\end{split}
\end{align}
If we make the natural choice of $s=1$, $q=1/C$, and $f=1/Y_{H}k$, then we are left with
\begin{align}
\begin{split}
\dot{F} &= \lambda F + \xi \rho RH - \sigma \left(1-R\right)F,  \\
\dot{H} &= \sigma \left(1-R\right)F - \xi \rho RH - \mu H,  \\
\dot{R} &= \alpha R\left(1-R\right) -\left(\rho R+\delta\right)H-\beta F
\end{split}
\end{align}
where we have dropped the subscripts on $\lambda_{max}$ and $\rho_{max}$ for simplicity, and $\xi=C/k$, $\delta=Y_{H}kP_{H}/C$, and $\beta=Y_{H}k\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right)/C$. The above equations represent the system of equations presented in the main text.

\section*{Parameter Values and Estimates}

%\subsection*{Standard synthesis and metabolic parameters}

All of the parameter values employed in our model have either been directly measured in previous studies or can be estimated from combining several previous studies. Below we outline previous measurements and simple estimates of the parameters.

Metabolic rate has been generally reported to follow an exponent close to $\eta=0.75$ (e.g., Refs. \citenum{West:2001bv,moses2008rmo} and the supplement for Ref. \citenum{hou}). We make this assumption in the current paper, although alternate exponents, which are know to vary between roughly $0.25$ and $1.5$ for single species \cite{moses2008rmo}, could be easily incorporated into our framework, and this variation is effectively handled by the $20\%$ variations that we consider around mean trends. The exponent not only defines several scalings in our framework, but also the value of the metabolic normalization constant, $B_{0}$, given a set of data.  For mammals the metabolic normalization constant has been reported to vary between $0.018$ (W g$^{-0.75}$) and $0.047$ (W g$^{-0.75}$; Refs. \citenum{hou,West:2001bv}, where the former value represents basal metabolic rate and the latter represents the field metabolic rate. We employ the field metabolic rate for our NSM model which is appropriate for active mammals (Table 1).

An important feature of our framework is the starting size, $m_{0}$, of a mammal which adjusts the overall timescales for reproduction. This starting size is known to follow an allometric relationship with adult mass of the form $m_{0}=n_{0}M^{\upsilon}$ where estimates for the exponent range between $0.71$ and $0.94$ (see Ref. \citenum{peters1986ecological} for a review). We use $m_{0}=0.097M^{0.92}$ \cite{blueweiss1978relationships} which encompasses the widest range of body sizes \cite{peters1986ecological}.

The energy to synthesize a unit of biomass, $E_{m}$, has been reported to vary between $1800$ to $9500$ (J g$^{-1}$) (e.g. Refs. \citenum{West:2001bv,moses2008rmo,hou}) in mammals with a mean value across many taxonomic groups of $5,774$ (J g$^{-1}$) \cite{moses2008rmo}. The unit energy available during starvation, $E^{\prime}$, could range between $7000$ (J g$^{-1}$), the return of the total energy stored during ontogeny \cite{hou} to a biochemical upper bound of $E^{\prime}=36,000$ (J g$^{-1}$) for the energetics of palmitate \cite{stryer,hou}. For our calculations we use the measured value for bulk tissues of $7000$ which assumes that the energy stored during ontogeny is returned during starvation \cite{hou}.

For the scaling of body composition it has been shown that fat mass follows $M_{\rm fat}=f_{0}M^{\gamma}$, with measured  relationships following  $0.018M^{1.25}$ \cite{Dunbrack:1993ec}, $0.02M^{1.19}$ \cite{Lindstedt:1985hm}, and $0.026M^{1.14}$ \cite{Lindstedt:2002td}. We use the values from \cite{Lindstedt:1985hm} which falls in the middle of this range. Similarly, the muscle mass follows $M_{\rm musc}=u_{0}M^{\zeta}$ with $u_{0}=0.383$ and $\zeta=1.00$ \cite{Lindstedt:2002td}.

%We also connect the resource growth rate to the total metabolic rate of an organism. That is, we are interested in the relative rates of resource recovery and consumption by the total population. From \cite{allen2002global} the total resource use of a population with an individual body size of $M$ is given by $B_{pop}=0.00061x^{-0.03}$ (W m$^{-2}$). Considering an energy density of $18200$ (J g$^{-1}$) of grass \cite{estermann} and an NPP between and $1.59\times10^{-6}$ and $7.92\times10^{-5}$ (g s$^{-1}$ m$^{-2}$) would give a range of resource rates between  $0.029$ and $1.44$ (W m$^{-2}$). This gives a ratio of total resource consumption to supply rates between $0.00042$ and $0.021$, and we used a value of $0.002$ in our calculations and simulations.

Typically the value of $\xi=C/k$ should roughly be $2$. The value of $\rho$, $\lambda$, $\sigma$, and $\mu$ are all simple rates (note that we have not rescaled time in our non-dimensionalization) as defined in the maintext. Given that our model considers transitions over entire stages of ontogeny or nutritional states, the value of $Y$ must represent yields integrated over entire life stages. Given an energy density of $E_{d}=18200$ (J g$^{-1}$) for grass \cite{estermann} the maintenance value is given by $P_{F}=B_{0}M^{3/4}/ME_{d}$, and the yield for a full organism will be given by $Y_{F}=ME_{d}/B_{\lambda}$ (g individual $\cdot$ g grass $^{-1}$), where $B_{\lambda}$ is the lifetime energy use for reaching maturity given by
\begin{equation}
B_{\lambda}=\int_{0}^{t_{\lambda}}B_{0}m\left(t\right)^{\eta}dt.
\end{equation}
Similarly, the maintenance for hungry individuals is $P_{H}=B_{0}(\epsilon_{\sigma}M)^{3/4}/(\epsilon_{\sigma}M)E_{d}$, and the yield for hungry individuals (representing the cost on resources to return to the full state) is given by $Y_{H}=ME_{d}/B_{\rho}$ where
\begin{equation}
B_{\rho}=\int_{\tau\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)}^{t_{\lambda}}B_{0}m\left(t\right)^{\eta}dt.
\end{equation}
Taken together, these relationships allow us to calculate $\rho$, $\delta$, and $\beta$.

Finally, the value of $\alpha$ can be roughly estimated by the NPP divided by the corresponding biomass densities. From the data in Ref. \citenum{michaletz2014convergence} we estimate the value of $\alpha$ to range between $2.81\times10^{-10}$ (s$^{-1}$) and $2.19\times10^{-8}$ (s$^{-1}$) globally. It should be noted that the value of $\alpha$ sets the overall scale of the $F^{*}$ and $H^{*}$ steady states along with the $B_{tot}$ for each type, and as such, we use $\alpha$ as our fit parameter to match these steady states with the data from Damuth \cite{damuth1987interspecific}. We find that the best fit is $\alpha=9.45\times10^{-9}$ (s$^{-1}$) which compares well with the calculated range above. However, two points are important to note here: first, our framework predicts the overall scaling of $F^{*}$ and $H^{*}$ independently of $\alpha$ and this correctly matches data, and second, both the asymptotic behavior and slope of $F^{*}$ and $H^{*}$ are independent of $\alpha$, such that our prediction of the maximum mammal size does not depend on $\alpha$.

%For the growth rate $\lambda$ we consider the standard model of $\ln\left(\upsilon\right)/t_{\lambda}$

@@

 \begin{table}[h]
\caption{Parameter values for mammals}
\label{param}
    \begin{center}
    \small
     \begin{tabular}{ p{1.2cm} p{3.2cm} l p{2.2cm}|}
     \hline
     Parameter & Value & References  \\
     \hline
   $\eta$ & $3/4$  &  (e.g. \cite{West:2001bv,moses2008rmo,hou}) \\
   $E_{m}$ & $5774$ (J gram$^{-1}$)  &  \cite{moses2008rmo,West:2001bv,hou} \\
   $E_{m}^{\prime}$ & $7000$  & \cite{stryer,hou} \\
   $B_{0}$ & $0.047$ (W g$^{-0.75}$)    & \cite{hou}  \\
%   $a$ & $1.78\times10^{-6}$  \quad \quad \\
%   $\lambda_{0}$ & $3.39\times10^{-7}$ (s$^{-1}$ gram$^{1-\eta}$) \quad \quad \\
   $\gamma$ & $1.19$ & \cite{Lindstedt:1985hm} \\
   $f_{0}$ & $0.02$ & \cite{Lindstedt:1985hm}\\
   $\zeta$ & $1.00$  & \cite{Lindstedt:2002td} \\
   $u_{0}$ & $0.38$  & \cite{Lindstedt:2002td} \\

   \hline
    \end{tabular}
    \end{center}
   \end{table}

\subsection*{Rate equations for invaders with modified body mass}
We allow an invading subset of the resident population with mass $M$ to have an altered mass $M^\prime = M(1+\chi)$ where $\chi$ varies between $\chi_{\rm min} <0$ and $\chi_{\rm max}>0$, where $\chi<0$ denotes a leaner invader and $\chi > 0$ denotes an invader with additional reserves of body fat.
Importantly, we assume that the invading and resident individuals have the same proportion of non-fat tissues.
For the allowable values of $\chi$ the adjusted mass should exceed the amount of body fat, $1+\chi>\epsilon_{\sigma}$, and the adjusted time to reproduce must be positive, which given our solution for $\tau(\epsilon)$ (see main text), implies that $1-\epsilon_{\lambda}^{1-\eta}\left(1+\chi\right)^{1-\eta}>0$.
Together these conditions imply that  $\chi\in(-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$ where the upper bound approximately equals $0.05$.

@@

% \bibliography{aa_starving_supplement}
\begin{thebibliography}{10}

\bibitem{Kempes:2012hy}
Kempes CP, Dutkiewicz S, Follows MJ
\newblock (2012) {Growth, metabolic partitioning, and the size of
  microorganisms.}
\newblock \emph{PNAS} 109:495--500.

\bibitem{kempes2014morphological}
Kempes CP, Okegbe C, Mears-Clarke Z, Follows MJ, Dietrich LE
\newblock (2014) Morphological optimization for access to dual oxidants in
  biofilms.
\newblock \emph{Proceedings of the National Academy of Sciences} 111:208--213.

\bibitem{West:2001bv}
West GB, Brown JH, Enquist BJ
\newblock (2001) {A general model for ontogenetic growth}.
\newblock \emph{Nature} 413:628--631.

>>>>

\RequirePackage{lineno}
\documentclass[twocolumn,preprintnumbers,amsmath,amssymb,superscriptaddress]{revtex4}
%\usepackage[pdftex]{graphicx}

\usepackage{amsmath,amsfonts,amssymb}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{color}
\usepackage{float}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{mathtools}
\usepackage{stmaryrd}
\usepackage{anyfontsize}

%\usepackage{epstopdf}
%\usepackage{array}
%\usepackage{tabularx}
%\usepackage{multirow}
\usepackage{color}
%\usepackage{multibox}
%\usepackage{rotating}
%\usepackage{lineno}
%\usepackage[left]{lineno}
\usepackage[comma,sort&compress]{natbib}
\bibpunct{}{}{,}{s}{}{;}
%\usepackage{authblk}
%\usepackage{multicol}

\bibliographystyle{naturemag}

% \usepackage{bibunits}

\linenumbers
% \setlength\linenumbersep{3pt}

@@

\author{Justin D. Yeakel} \affiliation{School of Natural Sciences, University
  of California, Merced, Merced, CA 95340, USA}\affiliation{The Santa Fe Institute, 1399 Hyde Park Road, Santa Fe, NM 87501, USA}\affiliation{Contributed equally}\affiliation{Corresponding author: jdyeakel@gmail.com}

\author{Christopher P. Kempes} \affiliation{The Santa Fe Institute, 1399 Hyde
  Park Road, Santa Fe, NM 87501, USA}\affiliation{Contributed equally}

\author{Sidney Redner} \affiliation{The Santa Fe Institute, 1399 Hyde Park
  Road, Santa Fe, NM 87501, USA}\affiliation{Contributed equally}

\title{ Supporting Information for ``The dynamics of starvation and recovery''}%: Eco-evolutionary feedbacks}

@@

%%%%%%%%%%%%%%%%%%%%%%%%
% SUPPLEMENTARY MATERIAL
%%%%%%%%%%%%%%%%%%%%%%%%

%New additions in red text

\clearpage

{\bf Mechanisms of Starvation and Recovery}
Our overall goal is to understand the dynamics of starvation, recovery, reproduction, and resource competition, where our framework partitions starvation and reproduction into two classes of the consumer: a full class that is able to reproduce and a hungry class that experiences mortality at a given rate and is unable to reproduce. For the dynamics of growth, reproduction, and resource consumption, past efforts have combined the overall metabolic rate as dictated by body size with a growth rate that is dependent on resource abundance and, in turn, dictates resource consumption (see Refs. \citep{Kempes:2012hy,kempes2014morphological} for a brief review of this perspective). This approach has been used to understand a range of phenomena including a derivation of ontogenetic growth curves from a partitioning of metabolism into maintenance and biosynthesis (e.g. \citep{West:2001bv,moses2008rmo,hou,Kempes:2012hy}) and predictions for the steady-state resource abundance in communities of cells \citep{kempes2014morphological}. Here we leverage these mechanisms, combined with several additional concepts, to define our Nutritional State Model (NSM).

We consider the following generalized set of explicit dynamics for starvation, recovery, reproduction, and resource growth and consumption
\begin{align}
\begin{split}
\dot{F_{d}} &= \lambda_{max} F_{d} + \rho_{max}R_{d}H_{d}/k - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho_{max}R_{d} H_{d}/k - \mu H_{d},  \\
\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) -\\
& \left[\left(\frac{\rho_{max}R_{d}}{Y_{H}k}+P_{H}\right)H_{d}+\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right)F_{d}\right].
\end{split}
\end{align}
where each term has a mechanistic meaning that we detail below (we will denote the dimensional equations with $_{d}$ before introducing the non-dimensional form that is presented in the main text). In the above equations $Y$ represents the yield coefficient (e.g., Refs. \citep{pirt,Heijnen}) which is the quantity of resources required to build a unit of organism (gram of mammal produced per gram of resource consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism). If we pick $F_{d}$ and $H_{d}$ to have units of (g organisms $\cdot$ m$^{-2}$), then all of the terms of $\dot{R_{d}}$, such as $\frac{\rho\left(R_{d}\right)}{Y}H_{d}$, have units of (g resource $\cdot$ m$^{-2}$ $\cdot$ s$^{-1}$) which are the units of net primary productivity (NPP), a natural choice for $\dot{R_{d}}$. This choice also gives $R_{d}$ as (g $\cdot$ m$^{-2}$) which is also a natural unit and is simply the biomass density. In this system of units $\alpha$ (s$^{-1}$) is the specific growth rate of $R_{d}$, and $C$ is the carrying capacity, or maximum density, of $R_{d}$ in a particular environment, and $k$ is the half-saturation constant (half the density of resources that would lead to maximum growth).
%(note this is not fully explicit because I don't know how to deal with the response of $\sigma$ to resources, although I have an idea for a derivation which may be necessary given the following approximations)

@@

We can formally non-dimensionalize this system by choosing the general rescaling of $F=fF_{d}$, $H=fH_{d}$, $R=qR_{d}$, $t=st_{d}$, in which case our system of equations becomes
%(ignoring the $\sigma (1-R)F$ terms which I don't have a dimensional form for yet),:
\begin{align}
\begin{split}
\dot{F} &= \frac{1}{s}\left[\lambda_{max} F + \rho_{max}\frac{R}{qk}H - \sigma \left(1-\frac{R}{qC}\right)F\right],  \\
\dot{H} &= \frac{1}{s}\left[\sigma \left(1-\frac{R}{qC}\right)F - \rho_{max}\frac{R}{qk} H - \mu H\right],  \\
\dot{R} &= \frac{1}{s}\left[\alpha R\left(1-\frac{R}{qC}\right) -\frac{q}{f}\left[\left(\frac{\rho_{max}R}{Y_{H}kq}+P_{H}\right)H+\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right) F\right]\right].
\end{split}
\end{align}
If we make the natural choice of $s=1$, $q=1/C$, and $f=1/Y_{H}k$, then we are left with
\begin{align}
\begin{split}
\dot{F} &= \lambda F + \xi \rho RH - \sigma \left(1-R\right)F,  \\
\dot{H} &= \sigma \left(1-R\right)F - \xi \rho RH - \mu H,  \\
\dot{R} &= \alpha R\left(1-R\right) -\left(\rho R+\delta\right)H-\beta F
\end{split}
\end{align}
where we have dropped the subscripts on $\lambda_{max}$ and $\rho_{max}$ for simplicity, and $\xi=C/k$, $\delta=Y_{H}kP_{H}/C$, and $\beta=Y_{H}k\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right)/C$. The above equations represent the system of equations presented in the main text.
\\

{\bf Parameter Values and Estimates}
All of the parameter values employed in our model have either been directly measured in previous studies or can be estimated from combining several previous studies. Below we outline previous measurements and simple estimates of the parameters.

Metabolic rate has been generally reported to follow an exponent close to $\eta=0.75$ (e.g., Refs. \citep{West:2001bv,moses2008rmo} and the supplement for Ref. \citep{hou}). We make this assumption in the current paper, although alternate exponents, which are known to vary between roughly $0.25$ and $1.5$ for single species \citep{moses2008rmo}, could be easily incorporated into our framework, and this variation is effectively handled by the $20\%$ variations that we consider around mean trends. The exponent not only defines several scalings in our framework, but also the value of the metabolic normalization constant, $B_{0}$, given a set of data.  For mammals the metabolic normalization constant has been reported to vary between $0.018$ (W g$^{-0.75}$) and $0.047$ (W g$^{-0.75}$; Refs. \citep{hou,West:2001bv}, where the former value represents basal metabolic rate and the latter represents the field metabolic rate. We employ the field metabolic rate for our NSM model which is appropriate for active mammals (Table 1).

An important feature of our framework is the starting size, $m_{0}$, of a mammal which adjusts the overall timescales for reproduction. This starting size is known to follow an allometric relationship with adult mass of the form $m_{0}=n_{0}M^{\upsilon}$ where estimates for the exponent range between $0.71$ and $0.94$ (see Ref. \citep{peters1986ecological} for a review). We use $m_{0}=0.097M^{0.92}$ \citep{blueweiss1978relationships} which encompasses the widest range of body sizes \citep{peters1986ecological}.

The energy to synthesize a unit of biomass, $E_{m}$, has been reported to vary between $1800$ to $9500$ (J g$^{-1}$) (e.g. Refs. \citep{West:2001bv,moses2008rmo,hou}) in mammals with a mean value across many taxonomic groups of $5,774$ (J g$^{-1}$) \citep{moses2008rmo}. The unit energy available during starvation, $E^{\prime}$, could range between $7000$ (J g$^{-1}$), the return of the total energy stored during ontogeny \citep{hou} to a biochemical upper bound of $E^{\prime}=36,000$ (J g$^{-1}$) for the energetics of palmitate \citep{stryer,hou}. For our calculations we use the measured value for bulk tissues of $7000$ which assumes that the energy stored during ontogeny is returned during starvation \citep{hou}.

For the scaling of body composition it has been shown that fat mass follows $M_{\rm fat}=f_{0}M^{\gamma}$, with measured  relationships following  $0.018M^{1.25}$ \citep{Dunbrack:1993ec}, $0.02M^{1.19}$ \citep{Lindstedt:1985hm}, and $0.026M^{1.14}$ \citep{Lindstedt:2002td}. We use the values from \citep{Lindstedt:1985hm} which falls in the middle of this range. Similarly, the muscle mass follows $M_{\rm musc}=u_{0}M^{\zeta}$ with $u_{0}=0.383$ and $\zeta=1.00$ \citep{Lindstedt:2002td}.

%We also connect the resource growth rate to the total metabolic rate of an organism. That is, we are interested in the relative rates of resource recovery and consumption by the total population. From \citep{allen2002global} the total resource use of a population with an individual body size of $M$ is given by $B_{pop}=0.00061x^{-0.03}$ (W m$^{-2}$). Considering an energy density of $18200$ (J g$^{-1}$) of grass \citep{estermann} and an NPP between and $1.59\times10^{-6}$ and $7.92\times10^{-5}$ (g s$^{-1}$ m$^{-2}$) would give a range of resource rates between  $0.029$ and $1.44$ (W m$^{-2}$). This gives a ratio of total resource consumption to supply rates between $0.00042$ and $0.021$, and we used a value of $0.002$ in our calculations and simulations.

Typically the value of $\xi=C/k$ should roughly be $2$. The value of $\rho$, $\lambda$, $\sigma$, and $\mu$ are all simple rates (note that we have not rescaled time in our non-dimensionalization) as defined in the maintext. Given that our model considers transitions over entire stages of ontogeny or nutritional states, the value of $Y$ must represent yields integrated over entire life stages. Given an energy density of $E_{d}=18200$ (J g$^{-1}$) for grass \citep{estermann} the maintenance value is given by $P_{F}=B_{0}M^{3/4}/ME_{d}$, and the yield for a full organism will be given by $Y_{F}=ME_{d}/B_{\lambda}$ (g individual $\cdot$ g grass $^{-1}$), where $B_{\lambda}$ is the lifetime energy use for reaching maturity given by
\begin{equation}
B_{\lambda}=\int_{0}^{t_{\lambda}}B_{0}m\left(t\right)^{\eta}dt.
\end{equation}
Similarly, the maintenance for hungry individuals is $P_{H}=B_{0}(\epsilon_{\sigma}M)^{3/4}/(\epsilon_{\sigma}M)E_{d}$, and the yield for hungry individuals (representing the cost on resources to return to the full state) is given by $Y_{H}=ME_{d}/B_{\rho}$ where
\begin{equation}
B_{\rho}=\int_{\tau\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)}^{t_{\lambda}}B_{0}m\left(t\right)^{\eta}dt.
\end{equation}
Taken together, these relationships allow us to calculate $\rho$, $\delta$, and $\beta$.

Finally, the value of $\alpha$ can be roughly estimated by the NPP divided by the corresponding biomass densities. From the data in Ref. \citep{michaletz2014convergence} we estimate the value of $\alpha$ to range between $2.81\times10^{-10}$ (s$^{-1}$) and $2.19\times10^{-8}$ (s$^{-1}$) globally. It should be noted that the value of $\alpha$ sets the overall scale of the $F^{*}$ and $H^{*}$ steady states along with the $B_{tot}$ for each type, and as such, we use $\alpha$ as our fit parameter to match these steady states with the data from Damuth \citep{damuth1987interspecific}. We find that the best fit is $\alpha=9.45\times10^{-9}$ (s$^{-1}$) which compares well with the calculated range above. However, two points are important to note here: first, our framework predicts the overall scaling of $F^{*}$ and $H^{*}$ independently of $\alpha$ and this correctly matches data, and second, both the asymptotic behavior and slope of $F^{*}$ and $H^{*}$ are independent of $\alpha$, such that our prediction of the maximum mammal size does not depend on $\alpha$.
\\
%For the growth rate $\lambda$ we consider the standard model of $\ln\left(\upsilon\right)/t_{\lambda}$

@@

 \begin{table}[h]
  {\color{red}
\caption{Parameter values for mammals}
\label{param}
    \begin{center}
    \footnotesize
     \begin{tabular}{p{3.8cm} c p{2.2cm} p{1.4cm}}
     \hline

     Definition & Parameter & Value & References  \\
     \hline
   Asymptotic adult mass & $M$ & (g) &  \\
   Initial mass of an organism & $m_{0}$ & (g) &  \\
   Metabolic rate scaling exponent & $\eta$ & $3/4$  &  (e.g. \citep{West:2001bv,moses2008rmo,hou}) \\
   Metabolic Normalization Constant & $B_{0}$ & $0.047$ (W g$^{-0.75}$)    & \citep{hou}  \\
   Initial mass scaling exponent & $\upsilon$ & $0.92$ &  \citep{blueweiss1978relationships,peters1986ecological} \\
   Initial mass scaling normalization constant & $n_{0}$ & $0.097$ (g$^{1-\upsilon}$) & \citep{blueweiss1978relationships,peters1986ecological}  \\   
   Fat mass scaling exponent & $\gamma$ & $1.19$ & \citep{Lindstedt:1985hm} \\
   Fat scaling normalization constant & $f_{0}$ & $0.02$ (g$^{1-\eta}$) & \citep{Lindstedt:1985hm}\\
   Muscle mass scaling exponent & $\zeta$ & $1.00$  & \citep{Lindstedt:2002td} \\
   Muscle scaling normalization constantv& $u_{0}$ & $0.38$ (g$^{1-\zeta}$)  & \citep{Lindstedt:2002td} \\
   Energy to synthesis a unit of mass & $E_{m}$ & $5774$ (J gram$^{-1}$)  &  \citep{moses2008rmo,West:2001bv,hou} \\
   Energy to synthesis a unit of mass during recovery & $E_{m}^{\prime}$ & $7000$ (J gram$^{-1}$) & \citep{stryer,hou} \\
   Specific resource growth rate & $\alpha$ & $9.45\times10^{-9}$ (s$^{-1}$) & see text  \\
   Fraction of asymptotic mass representing full state & $\epsilon_{\lambda}$ & $0.95$ & \citep{West:2001bv}  \\
   Fraction of asymptotic mass representing starving state & $\epsilon_{\sigma}$ & $1-f_{0}M^{\gamma-1}$ & see text  \\
   Fraction of asymptotic mass representing death & $\epsilon_{\mu}$ & $1-\frac{f_{0}M^{\gamma}+u_{0}M^{\zeta}}{M}$ & see text \\
   Carrying capacity (maximum density) of resources & $C$ & (g m$^{-2}$) & \\
   Half Saturation Constant & $k$ & (g m$^{-2}$) &   \\
   Normalized carrying capacity & $\xi$ & $C/k\approx2$ &   \\
   Reproductive fecundity & $\nu$ & $2$ & \citep{}  \\

%   $a$ & $1.78\times10^{-6}$  \quad \quad \\
%   $\lambda_{0}$ & $3.39\times10^{-7}$ (s$^{-1}$ gram$^{1-\eta}$) \quad \quad \\

   \hline
    \end{tabular}
    \end{center}
    }
   \end{table}

{\bf Rate equations for invaders with modified body mass}
We allow an invading subset of the resident population with mass $M$ to have an altered mass $M^\prime = M(1+\chi)$ where $\chi$ varies between $\chi_{\rm min} <0$ and $\chi_{\rm max}>0$, where $\chi<0$ denotes a leaner invader and $\chi > 0$ denotes an invader with additional reserves of body fat.
Importantly, we assume that the invading and resident individuals have the same proportion of non-fat tissues.
For the allowable values of $\chi$ the adjusted mass should exceed the amount of body fat, $1+\chi>\epsilon_{\sigma}$, and the adjusted time to reproduce must be positive, which given our solution for $\tau(\epsilon)$ (see main text), implies that $1-\epsilon_{\lambda}^{1-\eta}\left(1+\chi\right)^{1-\eta}>0$.
Together these conditions imply that  $\chi\in(-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$ where the upper bound approximately equals $0.05$.

@@

\bibliography{aa_starving_supplement}

% \end{bibunit}
